---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

# Using `{DT}` and `{githubinstall}` R packages to fetch and display in an interactive table github hosted R packages containing the term `phylogenetic` in teh repo description

# Loading libraries 

```{r}
library(githubinstall)
library(DT)
library(skimr)
library(BiocManager)
```


# Searching for packages relevant to `phylogenetic`

The following brief tutorial will guide you through the functions of the `{githubinstall}` R package, a package to easily query GitHub and find all relevant R packages to your selected query search term of interest. The function `githubinstall::gh_search_packages()` returns the list of R packages on GitHub that the titles contains a given #keyword .


For example, if you want to search packages that are relevant to `phylogenetic`, run the following:


```{r}
#  Search Packages by a Keyword, in this example we search for "phylogenetic" 
# install.packages("githubinstall") || found here: hoxo-m.github.io/githubinstall/

keyword <- "phylogenetic"

library(githubinstall)

pckges    <- githubinstall::gh_search_packages(keyword)
skimr::skim(pckges)
```

## Let's inspect what packages we retrieved related to `{phylogenetic}`

```{r}
DT::datatable(pckges)
```

Above we can see how many packages were found that contain the search term in their description. The `githubinstall::gh_search_packages` function returns 3 arguments for each package:

- `username`:account name of package developer on GitHub
- `package_name`:the name of the package that you can use as argument in the `install.packages()` function
- `title`: the description of the package

# 🤔 Mmm, wouldn't it be nice if the package name was actually a <clickable> link..

Let's try to add some interactivity to our retrieved table and convert the package names to clickable links that will take us to the github repo of the packages.

## Reconstructing the R package GitHub repo url for clickable links 

We will recontruct the string with the above pattern for all found packages, and add the html tag for redirecting so the pattern to recreate should look like this:

```{r}
# Mutate the package names to create clickable links:

pckges$package_name <- paste0("<a href='https://github.com/",
                             pckges$username, 
                             "/",
                             pckges$package_name, 
                             "'/",
                             "target='blank",
                             "'>",
                             pckges$package_name, 
                             "</a>")
```

## Updated datatable with clickable links!

Now we can inspect the packages that seem interesting by clicking the link. This will take us to the respective repository in GitHub.


```{r}
DT::datatable(pckges[,-1],      # the -1 to omit the username in the output table
              escape = FALSE   # this is required for the url
             )
```

